---
title: "Business Analytics with R Group Project"
authors: Yukyung Cha, Ismael Isak, Snehal Khatpe, Tashfia Mamun, and Naomi Snider
output:
  html_document:
    df_print: paged
---
Libraries
```{r Libraries}
#Load Libaries
library(ggplot2)
library(tidyr)
library(leaps)
library(glmnet)
library(RColorBrewer)
#install.packages("viridis")
library("viridis") 
library(neuralnet)
library(rpart)
library(caTools)
library(rpart.plot)
library(Amelia)
library(ggthemes)
library(corrplot)
library(corrgram)
require(quantmod)
require(PerformanceAnalytics)
require(xts)
library(tidyverse)
```
 
```{r Load Data Sets}
#Description of each company with division on sectors
security <- read.csv('C:/Users/littl/Downloads/securities.csv')
security
 
#Prices is raw as-is daily prices which does not account for stock splits
prices <- read.csv('C:/Users/littl/Downloads/prices.csv')
prices
 
#Split_prices is the same as the prices data, but has been adjusted for stock splits
split_prices <- read.csv('C:/Users/littl/Downloads/prices-split-adjusted.csv')
 
 
#converting dates
split_prices$date <- as.Date(split_prices$date)
#adding year to merge with fundamentals
split_prices$year <- format(split_prices$date , '%Y')
split_prices
 
#Metrics from SEC 10K filings
fundamentals <- read.csv('C:/Users/littl/Downloads/fundamentals.csv')
#converting dates
fundamentals$Period.Ending <- as.Date(fundamentals$Period.Ending)
#adding year to merge with split prices
fundamentals$year <- format(fundamentals$Period.Ending, '%Y')
fundamentals
```
 
Data Cleaning
```{r Data Cleaning}
#Merging the data sets to include split prices and fundamentals by year and symbol with all items from split prices being included
subset_selection <- merge(split_prices, fundamentals, by.x = c('symbol', 'year'), by.y = c('Ticker.Symbol','year'), all.x = TRUE)
#removing any NA values
subset_selection <- na.omit(subset_selection)
subset_selection <- subset(subset_selection, select = c('symbol','year', 'date', 'open', 'close', 'low', 'high', 'volume', 'After.Tax.ROE', 'Deferred.Asset.Charges', 'Deferred.Liability.Charges', 'Inventory', 'Investments', 'Net.Income', 'Operating.Income', 'Operating.Margin', 'Pre.Tax.ROE', 'Profit.Margin', 'Quick.Ratio', 'Total.Assets', 'Total.Equity', 'Total.Liabilities', 'Total.Revenue', 'Earnings.Per.Share', 'Estimated.Shares.Outstanding'))
 
subset_selection
```
 
```{r Descriptive Statistics}
## Taking the security file and dropping all the columns except GICS.Sector and Ticker.Symbol
symbol_sector <- security %>% select(Ticker.symbol, GICS.Sector)
symbol_sector
 
## Using left_join from dplyr to merge prices with symbol_sector based on stock symbol
names(symbol_sector)[1] <- "symbol"
new_prices <- left_join(split_prices, symbol_sector, by="symbol")
new_prices
 
#Mean opening price by industry for each year 
# group by industry (mean opening price)
tb <- new_prices %>%
  group_by(GICS.Sector, year) %>% 
  summarise(mean_open = mean(open))
 
ggplot(tb, aes(x=year, y=mean_open, color = GICS.Sector)) + geom_point()
ggplot(tb, aes(x=year, y=mean_open, color = GICS.Sector, group =1)) + geom_line()
ggplot(tb, aes(x=GICS.Sector,y=mean_open, fill = year)) + geom_bar(position = "dodge",stat = "identity")
tb
 
# group by industry (mean closing price)
tb1 <- new_prices %>%
  group_by(GICS.Sector, year) %>% 
  summarise(mean_close = mean(close))
 
ggplot(tb1, aes(x=year, y=mean_close, color = GICS.Sector)) + geom_point()
ggplot(tb1, aes(x=year, y=mean_close, color = GICS.Sector, group =1)) + geom_line()
ggplot(tb1, aes(x=GICS.Sector,y=mean_close, fill = year)) + geom_bar(position = "dodge",stat = "identity")
tb1
 
#boxplot of closing price by industry per year
ggplot(new_prices, aes(x=year, y=close)) + geom_boxplot() + facet_wrap(~ GICS.Sector)
 
#highest closing prices by company in each industry
tb3 <-new_prices %>%
  group_by(year, GICS.Sector, symbol,) %>% 
  summarise(mean_close = mean(close)) %>%
  arrange(desc(mean_close)) %>%
  slice(1:3)
 
ggplot(tb3, aes(x=year, y=mean_close, fill=symbol)) + 
  geom_bar(position = "dodge",stat = "identity") +facet_wrap(~ GICS.Sector) 
tb3
 
#We can see that top3 has remained pretty consistent across each industry
#Security – maybe a histogram of the industries 
table(security$GICS.Sector)
ggplot(security,aes(GICS.Sector)) + geom_bar() 
 
```
 
 
```{r Data Visualization}
#Data Visualization
missmap(prices, main="Prices Check", 
        col=c("yellow", "black"), legend=FALSE)
 
#security
missmap(security, main="Security Check", 
        col=c("yellow", "black"), legend=FALSE)
 
#split_prices
missmap(split_prices, main="split_prices Check", 
        col=c("yellow", "black"), legend=FALSE)
 
#Stock price per industry
symbol_sector <- security %>% select(Ticker.symbol, GICS.Sector)
prices$date <- as.Date(prices$date, "%Y-%m-%d")
names(symbol_sector)[1] <- "symbol"
new_prices <- left_join(prices, symbol_sector, by="symbol")
new_prices <- na.omit(new_prices)
 
plot1 <- ggplot(new_prices, aes(x=date, y=close, color=GICS.Sector)) + 
        geom_line() + theme_bw() + 
        scale_color_viridis(discrete = TRUE, option = "D") +
        scale_fill_viridis(discrete = TRUE, option = "D") +
        scale_x_date(breaks = '1 year')
plot1
 
#Stock price per subindustry
names(security)
unique(security$GICS.Sub.Industry)
symbol_sector <- security %>% select(Ticker.symbol, GICS.Sub.Industry)
prices$date <- as.Date(prices$date, "%Y-%m-%d")
names(symbol_sector)[1] <- "symbol"
new_prices <- left_join(prices, symbol_sector, by="symbol")
new_prices <- na.omit(new_prices)
 
plot2 <- ggplot(new_prices, aes(x=date, y=close, color=GICS.Sub.Industry)) + 
  geom_line() + theme_bw() + 
  scale_color_viridis(discrete = TRUE, option = "D") +
  scale_fill_viridis(discrete = TRUE, option = "D") +
  scale_x_date(breaks = '1 year')
plot2
 
# Relationship between close & After.Tax.ROE
names(subset_selection)
subset_selection$date <- as.Date(subset_selection$date, "%Y-%m-%d")
 
symbol_sector <- security %>% select(Ticker.symbol, GICS.Sector)
names(symbol_sector)[1] <- "symbol"
 
new_subset_selection <- left_join(subset_selection, symbol_sector, by="symbol")
new_subset_selection$date <- as.Date(new_subset_selection$date, "%Y-%m-%d")
 
plot3 <- ggplot(new_subset_selection, aes(x=After.Tax.ROE, y=close, color=GICS.Sector)) + 
  geom_point() +
  scale_color_viridis(discrete = TRUE, option = "D") +
  scale_fill_viridis(discrete = TRUE, option = "D")
plot3
 
#Zoom In
plot4 <- ggplot(new_subset_selection, aes(x=After.Tax.ROE, y=close, color=GICS.Sector)) + 
  geom_point() + scale_x_continuous(limits = c(0, 500)) +
  scale_color_viridis(discrete = TRUE, option = "D") +
  scale_fill_viridis(discrete = TRUE, option = "D")
plot4
 
# Relationship between close & Investments
summary(new_subset_selection$Investments)
 
plot5 <- ggplot(new_subset_selection, aes(x=Investments, y=close, color=GICS.Sector)) + 
  geom_point() +
  scale_color_viridis(discrete = TRUE, option = "D") +
  scale_fill_viridis(discrete = TRUE, option = "D")
plot5
 
#Stock price per industry
symbol_sector <- security %>% select(Ticker.symbol, GICS.Sector)
prices$date <- as.Date(prices$date, "%Y-%m-%d")
names(symbol_sector)[1] <- "symbol"
new_prices <- left_join(prices, symbol_sector, by="symbol")
new_prices <- na.omit(new_prices)
 
# Flow of difference per industry
new_subset_selection$date <- as.Date(new_subset_selection$date, "%Y-%m-%d")
plot6 <- ggplot(new_subset_selection, aes(close, fill = GICS.Sector)) + 
  geom_histogram(position="dodge") +
  scale_color_viridis(discrete = TRUE, option = "D") +
  scale_fill_viridis(discrete = TRUE, option = "D")
plot6
 
plot7 <- ggplot(new_subset_selection, aes(close, fill = GICS.Sector)) + 
  geom_boxplot()+ 
  scale_color_viridis(discrete = TRUE, option = "D") +
  scale_fill_viridis(discrete = TRUE, option = "D")
plot7
 
plot8 <- ggplot(new_subset_selection, aes(x=GICS.Sector, y=close, color = GICS.Sector)) + 
  geom_point() +
  scale_color_viridis(discrete = TRUE, option = "D") +
  scale_fill_viridis(discrete = TRUE, option = "D")
plot8
 
```
 
```{r Preparing the Data}
# ----FORWARD STEPWISE----
fwd_fit <- regsubsets(close~., subset_selection, nvmax = 20, method="forward")
fwd_summ <- summary(fwd_fit)
fwd_summ
 
# using the adj R squared to check goodness of fit
fwd_summ$adjr2
which.max(fwd_summ$adjr2)
 
#getting the coefficients
coef(fwd_fit, 15)
 
 
#----BACKWARD STEPWISE----
back_fit <- regsubsets(close~., subset_selection, nvmax = 20, method="backward")
back_summ <- summary(back_fit)
back_summ
 
### get the adjusted R squared
back_summ$adjr2
which.max(back_summ$adjr2)
coef(back_fit, 15)
 
```
 
 

```{r Evaluation Algorithms}
#----RIDGE REGRESSION----
subset_selection_2 <-  subset(subset_selection, select = c('open', 'close', 'low', 'high', 'volume'))
x <- model.matrix(close~., subset_selection_2)[,-1]  # matrix of the predictors values
y <- subset_selection_2$close  # vector of the dependent variable values
w <- seq(10, -3, length=100)
lvalues <- 10 ^ w
lvalues

#Fit the ridge regression model
ridge_fit <- glmnet(x, y, alpha = 0, lambda = lvalues)
ridge_fit$lambda

#Lambda for the model
ridge_fit$lambda[40]
best_lambda <- ridge_fit$lambda.min

#Coefficient for the model
coef(ridge_fit)[,40]
predict(ridge_fit, s = 1200, type = "coefficients")

#Improved lambda
best_ridge_fit <- glmnet(x, y, alpha = 0, lambda = best_lambda)
best_ridge_fit$lambda
coef(best_ridge_fit)

#prediction from improved model
y_predicted <- predict(best_ridge_fit, s=best_lambda, newx = x)
 

#find SST and SSE
sst <- sum((y - mean(y))^2)
sse <- sum((y_predicted - y)^2)
sse

#find R-Squared
rsq <- 1 - sse/sst
rsq

rsq

plot(ridge_fit)
plot(ridge_fit, xvar = "lambda")


#Mean Squared Error
mse <- sse/(324240-3)
mse

#----LASSO REGRESSION----
#define response variable
y <- subset_selection$close
#define matrix of predictor variables
x <- data.matrix(subset_selection[, c('open', 'close', 'low', 'high', 'volume')])


#perform k-fold cross-validation to find optimal lambda value
cv_model <- cv.glmnet(x, y, alpha = 1)

#find optimal lambda value that minimizes test MSE
best_lambda <- cv_model$lambda.min
best_lambda

#find coefficients of best model
best_model <- glmnet(x, y, alpha = 1, lambda = best_lambda)
coef(best_model)
y_predicted <- predict(best_model, s = best_lambda, newx = x)
#find SST and SSE
sst <- sum((y - mean(y))^2)
sse <- sum((y_predicted - y)^2)
sse
#find R-Squared
rsq <- 1 - sse/sst
rsq
#find MSE
lasso_MSE <- sse/(324240-3)
lasso_MSE

plot(cv_model)



#----DECISION TREE----
head(prices)
### create the training set and the test set 

i <- sample(342240, 162120)
SE_train <- subset_selection_2[i, 1:5]
SE_test <- subset_selection_2[-i, 1:5] 

### grow the regression tree with the rpart() function
### the method parameter must be set to "anova" 

### the rpart() function has built-in cross validation
### it performs a 10-fold cross-validation
fit <- rpart(close~., data = SE_train, method = "anova") 

### plot the tree with the prp() function
prp(fit) 

### plot the tree with the rpart.plot() function
rpart.plot(fit) 

### print the complexity parameter table
printcp(fit) 


### compute the goodness-of-fit in the TEST set
pred1 <- predict(fit, SE_test)
mse <- sum((pred1 - SE_test$close)^2)/162120
mse 

var.y <- sum((SE_test$close - mean(SE_test$close))^2)/162119 

rsq <- 1 - mse/var.y
rsq
 
#----NEURAL NETWORK----
##Neural Network to predict stock prices
 
#splitting the data
split_nn = sample.split(subset_selection_2$close, SplitRatio = 0.70)
 
train_nn = subset(subset_selection_2, split_nn == TRUE)
test_nn = subset(subset_selection_2, split_nn == FALSE)
 
#Training the Model to predict closing price
 
 
nn <- neuralnet(close ~ open + low + high + volume
                ,data=train_nn,hidden=2,linear.output=TRUE)
 
#neuralnet visualization
plot(nn)
 
# Compute Predictions off Test Set
predicted.nn.values <- neuralnet::compute(nn,test_nn[c('open','low','high','volume')])
 
# Check the Mean Squared Error
MSE.nn <- sum((test_nn$close - predicted.nn.values$net.result)^2)/nrow(test_nn)
MSE.nn
 
#find SST and SSE
sst_nn <- sum((test_nn$close - mean(test_nn$close))^2)
sse_nn <- sum((predicted.nn.values$net.result - test_nn$close)^2)
#find R-Squared
rsq_nn <- 1 - (sse_nn/sst_nn)
rsq_nn
 
 
 
#----LINEAR REGRESSION----
df3 <-subset_selection_2
df3 <- as.data.frame(df3)
 
head(df3)
num.cols3 <- sapply(df3, is.numeric)
cor.data3 <- cor(df3[,num.cols3])
cor.data3
#Corrplot
corrplot(cor.data3, method='color')
#Corrgram
corrgram(df3, order=TRUE, lower.panel=panel.shade,
         upper.panel=panel.pie, text.panel=panel.txt)
#Histogram
ggplot(df3,aes(x=close)) + geom_histogram(bins=20,alpha=0.5,fill='blue') + theme_minimal()
 
set.seed(101)
sample3 <- sample.split(df3$close, SplitRatio = 0.70)

train3 = subset(df3, sample3 == TRUE)
test3 = subset(df3, sample3 == FALSE)
 
model3 <- lm(close ~ open + low + high + volume , train3)
summary(model3)

lr_predictions <- predict(model3,test3)

results <- as.data.frame(cbind(lr_predictions, test3$close))
colnames(results) <- c('pred', 'real')

SSE = sum((results$pred - results$real)^2)
SST = sum( (mean(split_prices$close) - results$real)^2)

R2 = 1 - SSE/SST
R2

mse <- mean((results$real-results$pred)^2)
mse

#Visualize
res3 <- residuals(model3)
res3 <- as.data.frame(res3)
head(res3)
 
#Histogram of residuals
ggplot(res3, aes(res3)) + geom_histogram(fill='blue', alpha=0.5)

#Scatterplot of residuals
plot(fitted(model3), resid(model3), xlab="Fitted values", ylab="Residuals", col="black", pch = 16)
abline(0,0)
 
# Result Visualization
# Creating data frame for visualization (model vs real)



# Setting x value as index(similar to the length of each column) since it's impossible to bring back the date values after mixing up and sampling the data into train&test set
results_ggp <- data.frame(x = 1:92072,
                          y = c(results$pred, results$real),
                          group = c(rep("pred", nrow(results)),
                                    rep("real", nrow(results))))





result_viz <- ggplot(results_ggp, aes(x, y, col = group)) +
  geom_line() + xlab("Index") + ylab("Stock Price") +
  ggtitle("Comparison: Model Predictions vs Real Value") +
  scale_color_viridis(discrete = TRUE, option = "D") +
  scale_fill_viridis(discrete = TRUE, option = "D")
result_viz
 0
 
#Save model for later use
saveRDS(model3, file = "nyse_linear_regression_model.rda")
 
# loading the model
previous_model = readRDS("nyse_linear_regression_model.rda")

```



Best performing stock analysis:
First I'm going to create separate groups based on ticker symbol.

```{r}
individual_stocks <- split_prices %>% group_by(symbol) %>% nest()
```

Then I'm going to change the data type for each stock in to a Time Series format that we can use.

```{r}
for(i in 1:501){
  individual_stocks[[2]][[i]] <- xts(individual_stocks[[2]][[i]][,2:6], order.by=as.Date(individual_stocks[[2]][[i]][,1]$date))
}
```


## Finding best performing stocks

The criterea for best performing stocks would be the last closing price minus the first opening price divided by the last closing price.

$$
\frac{Last.Price-First.Price}{Last.Price}
$$

```{r}
ROI <- matrix(NA,nrow = 501, ncol = 1)
stock <- matrix(NA,nrow = 501, ncol = 1)
```
```{r}
for(i in 1:501){

  stock[i] <- individual_stocks[[1]][[i]]
}
```
```{r}
for(i in 1:501){
ROI[i] <- (as.numeric(last(individual_stocks[[2]][[i]][,2]))-as.numeric(first(individual_stocks[[2]][[i]][,1]))/as.numeric(last(individual_stocks[[2]][[i]][,2])))
}
```
```{r}
Price_change <- cbind(stock, ROI)

Price_change[order(Price_change[,2], decreasing = TRUE),][1:10,]
```

## Simple Trading strategy

Using those 10 best performing stocks we can perform a simple trading strategy where we buy at the open and sell at the close. Our trading signal would be then be whether the price increased by 5% since the last closing, this is so we can maximize our earning and avoid downturns in the market.

```{r}
NFLX <- individual_stocks %>% filter(symbol=="NFLX")



price <- Cl(NFLX$data[[1]]) # close price
r <- price/Lag(price) - 1 # % price change
delta <-0.005 #threshold
signal <-c(0) # first date has no signal

#Loop over all trading days (except the first)
for (i in 2: length(price)){
  if (r[i] > delta){
    signal[i]<- 1
  } else
    signal[i]<- 0

}
signal<-reclass(signal,price)


trade <- Lag(signal,1)
ret<-dailyReturn(NFLX$data[[1]])*trade
names(ret)<-"filter"
```


This chart shows us the performance of $1 based on the strategy of only trading only if the price increases by 5% from the previous closing.

```{r}
chart.CumReturns(ret, main="Simple Strategy", wealth.index = TRUE)
```

This would compare to no strategy in this chart.

```{r}
chartSeries(NFLX[[2]][[1]],
            show.grid = FALSE,
            TA = NULL,
            name = "NFLX",
            type = 'line',
            theme=chartTheme('white'))
```

This basically tells us that having even a basic strategy when investing yields better results than without because not all stocks perform as well as NFLX.
